Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-08 Ben_Manuscripts/structure_paper/Draft.tex
updates

  NF was introduced as an intermediate between RO and ultrafiltration, having
  the ability to separate organic matter and salts on the order of one nanometer
  in size. Explicit pore pathways running through the thickness of the membrane
  permit high solvent fluxes at lower pressures than RO. NF membranes typically
  have a surface charge resulting from ionizable groups which affords separation
  of ions smaller than the pore readius by Donnan exclusion.
  \cite{van_der_bruggen_review_2003} NF is often used as a precursor to reverse
  osmosis since it can efficiently remove a significant portion of dissolved
  solutes. Following NF pretreatment, RO can further purify the permeate at lower
  pressure. Unfortunately, NF membranes, like RO, possess a pore size
  distribution which limits their ability to perform precise separations
  \cite{bowen_modelling_2002}. An NF membrane with high selectivity could offer
  the versatility to separate ions as well as small charged and uncharged
  molecules to the same degree as RO and at a significantly lower cost.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer named
  Na-GA3C11 have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h} phase.
  The presence of ca. 10 wt \% water results in the H\textsubscript{II} phase. In
  both cases, monomers assemble into mesophases made of hexagonally packed,
  uniform size, cylinders with hydrophilic head groups oriented inward towards
  the cylinder center.  The hydrophilic region can act as a pore for aqueous
  separations.  One can envision tailoring the pore region for specific
  separations by changing the monomer chemsitry.

@@

  % BJC: Not sure this figure needs to be in the paper. 
  %MRS2: good question.  It helps give people context, which is necessary to
  %move forward.  Perhaps leave it in for now, though may need to get
  %permission to reprint it.
  \begin{figure} \centering
  \includegraphics[width=\linewidth]{soft_confinement.png} \caption{The wedge
	  shaped liquid crystal monomer (a) self assembles into mesophases with
		  hexagonally packed pores (b). The pores are made of stacked monomer disks. A
		  sub-micron-thick film is created by casting a dilute solution of Na-GA3C11/THF
		  solution onto a silicon substrate and and allowing the solvent to evaporate.
		  The thin film contains nanoporous columns which lie parallel to the film plane.
		  (d) When a soft PDMS pad is imposed to the thin film, with subsequent thermal
		  annealing, the columns align perpendicular to the film plane. (e)
		  Photo-cross-linking of the aligned film creates a mechanically stable thin film
		  with vertically aligned nanopores}~\label{fig:soft} \end{figure}

@@

  %BJC: The Methods sections has four main topics which will be separated by an appropriate header

  %%%%%%%%% Monomer Preparation %%%%%%%%%%%%%

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using the am1bccsym method of molcharge shipped with QUACPAC from Openeye % cite with webpage
  Scientific Software. All molecular dynamics simulations were run using Gromacs 2016  % This the version on Bridges
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned with molcharge. 
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

  %%%%%%%% UnitCell Preparation %%%%%%%%%%%%%

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system. 
  \begin{itemize}
    \item Previous work has shown a coarse grain model self assemble into
    the H\textsubscript{II} phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.
    \item We attempted atomistic self-assembly by packing monomers into a box 
    using Packmol \cite{martinez_packmol:_2009}.
    \item Simulations of greater than 100 ns show no indicators of progress 
    towards an ordered system.  %BJC: include more details in supplment
    \item To bypass the slow self-assembly process, python scripts are used
    to assemble monomers into a structure close to one of a number of 
    hypothesized equilibrium configurations (Figure~\ref{fig:python}).
  \end{itemize}

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
%    \item Each layer is made of a set number of monomers per layer, a variable to
%    be explored further.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    %MRS2: I was referring to the data choice of 20 vs 10.  Yes, you can refer to it here, because the expectation is that peple will only look at it after reading the whole paper.
%    \item We chose initial guesses for the remaining structural parameters based on
%    experimental data and treated them as variables during model development.
  \end{itemize}

%  We chose an initial layer spacing based on experimental 2D WAXS data.
%  \begin{itemize}  
%    \item The pattern shows reflections corresponding to features spaced 3.7 \angstrom~apart.
%    \item It has been hypothesized that the features are a result of $\pi$-$\pi$
%    interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
%    \item Our simulations tend to equilibrate to a wider interlayer spacing of
%    $\approx$ 4.1 \angstrom~, which inspired separate systems starting with layer 
%    spacings greater than 4 \angstrom~.
%  \end{itemize}

  %%%%%%%%%%%% Monomer Placement %%%%%%%%%%%%%%

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others.
  \begin{itemize}
	\item The starting pore radius, defined as the distance of a chosen 
        head group carbon from the pore's central axis, does not influence the
        equilibrium structure when a reasonable value is chosen (See Supplemental)
        \item The pore radius is chosen to be 0.6 nm in our initial configurations 
        because the pore size is estimated to be $\approx$ 1.2 nm % Add reference
	\item The initial distance between pores also has little effect on the 
        the equilibrated structure. However, one should not start them too close or there 
        will likely be unintended high energy repulsions during early equilibration
	\item We chose an initial pore spacing of 4.5 nm, $\approx$ 10 \% larger than 
        the experimental value of 4.12 nm.
        \item The distance between layers and the rotation of the layers with respect
        to adjacent layers, and the number of monomers per layer do influence the 
        equilibrium structure and require further justification for their choices. 
        \item We rely on experimental data to inform our choices
  \end{itemize}

  We chose the layer spacing for the initial configuration based on R-$\pi$ 
  and then allowed system to readjust during equilibration.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be coplanar with the xy plane.
    \item We explore two different initial layer spacings
    \item The first is exactly equal to R-$\pi$ with layers placed so aromatic
    rings are stacked 3.7 \angstrom~ apart in the z-direction.
    \item A second system is explored with an initial layer spacing of 5 \angstrom
%MRS3: at least state in the main body what went wrong with it in a sentence or so, with details in supplemental. 
   \item A third system with an initial layer spacing of 10 \angstrom was briefly
    explored
   \item When layers are spaced out sufficiently, they tend to collapse on each other 
   while simultaneously slipping in between layers of other pores which lead to an
   artificially thick membrane with pores spaced closely together.
   \item The intereseted reader can learn more about it in the Supplemental Information.
% BJC: Move all this lower down to be with the rest of the equilibrium calculations
%    \item To quantify the degree of layering in our system, we calculated
%    a spatial correlation function, $g(z)$, measured along the z-axis 
%    (perpendicular to the membrane plane)
%    \item To calculate $g(z)$, we binned the z component distances between 
%    the center of mass of each benzene ring and all others of the same pore 
%    over 50 ns of equilibrated trajectory and then normalized by the average
%    number density.
%    \item We compare the degree of layering between systems based on the
%    amplitude of the first peak in $g(z)$. We halve the difference between the
%    maximum of the first peak and the following minimum. We compare the difference
%    to the mean to get a percentage deviation from the average number density.
%    \item To extract the average distance between layers we applied a discrete 
%    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

% BJC: removing this since I say earlier that the system is insensitive to this choice
% Detail will be provided in the supplement
% discussion of measuring equilibrium pore spacing is moved down to calculations part of methods
%  We placed pores at a chosen initial spacing based on R-pores, then allowed
%  the system to settle into its preferred spacing. 
%  \begin{itemize}
    %MRS: somewhere, should explain there's a large dependence of (meta)stable structures based on this initial distance.  It's sort of important to put somewhere.
    %BJC: Sure, but different choices of pore spacing were based on number of monomers per layer. More monomers per layer means a larger initial pore spacing.
%    \item The model's pore centers are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
%    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
%    resulting from a tightly packed initial configuration.
%    \item To calculate the equilibrated pore spacing, we measured the distance between pore
%    centers.
%    \item Pore centers were located by averaging the coordinates of sodium ions in their 
%    respective pores.
%    \item Statistics were generated using the bootstrapping technique (See Supplemental Information)
    	% Supplemental vvvv
%	\begin{itemize}
        %MRS2: make clear sampling what?
%        \item We are interested in 5 pore-to-pore distances which should all be equal in
%        a perfect hexagonal array, however only 4 distances are independent % can visualize this in supplemental material. Will make a lot of things more clear below.
%	\item Each pore spacing has its own trajectory of spacing vs. time. Using data collected after
%	the system is equilibrated, we calculate how long it takes for the data in each of the 5
%	trajetories to become uncorrelated using pymbar.timeseries.integratedAutocorrelationTime() % reference
%	\item We break the full trajectories down into sub-trajectories based on the
%	maximum autocorrelation time of those found in the previous step. 
%    	\item For each bootstrap trial, we recreate an equilibrium trajectory by randomly 
%    	sampling pore spacings from the sub-trajectories 
%    	\item We get an average value for each pore spacing by finding the mean of the bootstrapped data
%	\item We calculate the overall average as the mean of all bootstrapped pore spacings
%	\item The uncertainty for each pore spacing is calculated as $\dfrac{<x> - x}{4}$ where $<x>$ is the
%	average spacing from the bootstrap trial and $x$ is the average value of one of the pore spacings.
%	\item We report the mean of these uncertainties
        % MRS: need to deal with the fact only 4 are independent in the bootstrapping?  How did we do that?
	% BJC: Updated
        % Supplemental ^^^^
%	\end{itemize}
%  \end{itemize}

% BJC: This can be covered more in the Supplmental ... it might not even be necessary. I won't
% report any "pore sizes" except in the supplemental where I will show that the initial choice
% of pore radius has little effect on pore size

%  We used experimental Transmission electron microscopy (TEM) and size exclusion rejection
%  data \cite{feng_scalable_2014,feng_thin_2016,zhou_supported_2005} to inform our definition
%  of pore radius in the initial configuration.
%  \begin{itemize}
%    \item Experimental evidence suggests uniform pores with radii of 0.6 nm 
%    \item Comparing a geometric measurement of pore size derived from an atomistic model,
%    to a less precise, experimentally derived pore size estimate, will give ambiguous results.
%    \item What is meant by pore radius will not be clear until we establish a clear picture
%    of the nanoscopic pore environment.
%    \item When constructing pores, we chose the carboxylate carbon from the monomer
%    head group as a reference atom, and placed it a distance r from the pore center,
%    where r is the pore radius (~\ref{fig:pore_radius_illustration}). % BJC: Figure in Supplemental Information?
%    \item We will not make direct comparisons of pore radius between our model 
%    and experiment to avoid the ambiguity, however, we do define a pore radius  
%    for our own purposes.
    %\item To measure the pore radius in our model we calculate the distance between 
    %the center of mass of each aromatic ring and the center of mass of all aromatic rings
    %in their respective pores.
%  \end{itemize}

%MRS2: is the pore radius of the initial configuration really relevant? Does it need to be defined?  Seems like what radius means in the final equilibrated one is more important.
%BJC: This can go in supplemental
%  \begin{figure}
% 	\centering
% 	\includegraphics[width=\textwidth]{pore_radius_illustration.png}
%	\caption{The pore radius of the initial configuration is defined using the 
%        carboxylate carbon as a reference}~\label{fig:pore_radius_illustration}
%  \end{figure}

  The relative interlayer orientation was chosen based on clues from 
  diffraction data as well as the various known stacking modes of benzene 
  and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  \begin{itemize}
    \item The T-shaped configuration was ruled out based on the inconsistency of
    its $\approx$ 5 \angstrom~equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}.
    \item The system's preference towards the sandwiched vs. parallel displaced 
    stacking modes will be explored.
    \item Both have reported stacking distances near the R-$\pi$ value of 3.7 \angstrom % Reference
    \item Headgroups in our sandwiched initial configuration are stacked directly
    on top of each other while stacked headgroups in the parallel displaced 
    initial configuration are offset by $180/nmon$ degrees where $nmon$ equals
    the number of monomers per layer.
    %BJC: I think the following commentary belongs elsewhere, maybe even in the discussion
    %MRS: hard to say. It's certainly a big switch from things that are relatively straightforward to things that require 
    % a fair amount of information, so I think it belongs in it's own section -- so far, you're just laying out the approach.  This is what we fix, this is what we leave as a variable.  The analysis probably belongs later.
    %BJC: Yeah, I will leave the rest for discussion
%     \item Visualization of each configuration (\Cref{fig:sandwichedlayers,fig:offsetlayers})
%     suggests entropic differences based on the way the tails are able to pack.
%     \item In the sandwiched configuration, all tails start out directly on top
%     of each other which may prevent closely stacked benzene rings.
%     \item In the offset configuration, the tails are placed in between each other 
%     which may allow layers to come together in a compact way.
%     \item This difference may explain, in part, which stacking mode is more favorable.
  \end{itemize}

  As outlined in \ref{point:monomernum} the number of monomers in each layer is unknown
  so we tested configurations constructed with a varied number of monomers per layer.
  \begin{itemize}
	\item Systems were built in the offset and parallel displaced configurations
        with 4, 5, 6, 7 and 8 monomers per layer
  \end{itemize}

  %%%%%%%%%%%% Equilibration %%%%%%%%%%%%%%

  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
        %MRS3: then just state what you did, so it could be repeated. 
          \item Restraints with a force constant of 1e6 KJ mol$^-1$ nm$^-2$, fix monomer 
	  head groups in the sandwiched or parallel-displaced configurations while 
          allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value %, starting from 1e6 KJ mol$^-1$ nm$^-2$. <-- stated in first bullet
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All water molecules placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}

%  An equilibration scheme with position restraints placed on aromatic rings
  %MRS: 
  %prevents 
%  during early equilibration steps is required to prevent high energy repulsions .
  %MRS: should be clearer what you mean by ``unrealistic jumps''.  Hard to visualize.
  %BJC: 'high energy repulsions' ?
%  \begin{itemize}
%    \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
%    configurations while allowing monomer tails to settle.
%    \item Doing so also mitigates system dependence on initial monomer configuration.
%    \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
%    \item Every 50 ps, we reduce the force constants by the square root of its 
%    previous value, starting from 1e6 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}.
%    \item Once the force constant is below 10 KJ mol\textsuperscript{-1} 
%    nm\textsuperscript{-2}, the restraints are slowly released until there is no more 
%    restraining potential.  
%    \item The resulting unrestrained structure is allowed to equilibrate further in the NPT
%    ensemble for 400 - 500 ns.
   %BJC: I will reference an equilibration script that can be released with the supplemental info
   %MRS: we should release as many scripts as is feasible, can be in a github repository.
   %BJC: Okay, once we submit, I will need to spend some time cleaning the scripts
   %MRS: You've described two equilibration procedures (or two varieties of the same one?) - need to be crystal clear to the reader as to which is used where.
   %BJC: Yea I'm just saying that I do 50 ps restrained simulations in NVT ensemble and then once the restraints are gone I switch to NPT. I edited the language to try and make it clearer.
%  \end{itemize}

  %BJC: the following paragraph will be replaced with something that Joe writes
  %MRS: sounds good.

  % BJC: The following methodology is taken from the 2014 Feng et al. paper where
  % WAXS is reported. The data we are using comes from the 2016 paper which 
  % I assume used the same set up but I will need to verify this with Xunda
  % before this gets sent out. The SAXS is the same as shown in the 2016 paper
  % however I will likely reproduce it with matplotlib

  %Experimental wide angle X-ray scattering measurements were performed using a 
  %Rigaku 007 HF+ instrument with a rotating anode Cu K$\alpha$ X-ray source
  %($\lambda$~= 1.542 \angstrom~) and a 2-D Saturn 994+ CCD detector. The 
  %calibrations of the resultant 2-D WAXS patterns were done by using a 
  %silicon powder standard (d-spacing of 3.1355 \angstrom~). 2-D SAXS scattering
  %patterns were integrated into 1-D plots of scattering intensity (I) versus
  %q, where q = 4$\pi$sin($\theta$)/$\lambda$ and the scattering angles is
  %2$\theta$. The data shown is reproduced from data collected elsewhere~\ref{feng_scalable_2014}.

%%%%%%%%%%%%% Crosslinking %%%%%%%%%%%%

>>>>

  NF was introduced as an intermediate between RO and ultrafiltration, having
  the ability to separate organic matter and salts on the order of one nanometer
  in size. Explicit pore pathways running through the thickness of the membrane
  permit high solvent fluxes at lower pressures than RO. NF membranes typically
  have a surface charge resulting from ionizable groups which affords separation
  of ions smaller than the pore readius by Donnan exclusion.
  \cite{van_der_bruggen_review_2003} NF is often used as a precursor to reverse
  osmosis since it can efficiently remove a significant portion of dissolved
  solutes. Following NF pretreatment, RO can further purify the permeate. 
  Unfortunately, NF membranes, like RO, possess a pore size
  distribution which limits their ability to perform precise separations
  \cite{bowen_modelling_2002}. An NF membrane with high selectivity could offer
  the versatility to separate ions as well as small charged and uncharged
  molecules to the same degree as RO and at a significantly lower cost.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer named
  Na-GA3C11 have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase \ref{fig:assembly} The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemsitry.

@@

  % BJC: Not sure this figure needs to be in the paper. 
  %MRS2: good question.  It helps give people context, which is necessary to
  %move forward.  Perhaps leave it in for now, though may need to get
  %permission to reprint it.
  %\begin{figure} \centering
  %\includegraphics[width=\linewidth]{soft_confinement.png} \caption{The wedge
  %	  shaped liquid crystal monomer (a) self assembles into mesophases with
%		  hexagonally packed pores (b). The pores are made of stacked monomer disks. A
%		  sub-micron-thick film is created by casting a dilute solution of Na-GA3C11/THF
%		  solution onto a silicon substrate and and allowing the solvent to evaporate.
%		  The thin film contains nanoporous columns which lie parallel to the film plane.
%		  (d) When a soft PDMS pad is imposed to the thin film, with subsequent thermal
%		  annealing, the columns align perpendicular to the film plane. (e)
%		  Photo-cross-linking of the aligned film creates a mechanically stable thin film
%		  with vertically aligned nanopores}~\label{fig:soft} \end{figure}

  %BJC: Made my own version of the above, leaving out alignment details since it'

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		%\includegraphics[width=\linewidth, trim={2.5cm 0 4cm 1.25cm}, clip]{WAXS_raw.png}
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnnar mesophases (e)}~\label{fig:assembly}
  \end{figure}

@@

  \subsection*{Monomer Placement}

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of molcharge shipped with QUACPAC from Openeye Scientific Software. All
  molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).

  \subsection*{Unit Cell Preparation}

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system.  Previous
  work has shown a coarse grain model self assemble into the H\textsubscript{II}
  phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (See supplemental information
  for further details). To bypass the slow self-assembly process, python scripts
  are used to assemble monomers into a structure close to one of a number of
  hypothesized equilibrium configurations (Figure~\ref{fig:python}).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. This point will be explained in more detail later.

\subsection*{Monomer Placement}

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm. The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will likely be unintended high energy repulsions during early
  equilibration. We chose an initial pore spacing of 4.5 nm, $\approx$ 10 \%
  larger than the experimental value of 4.12 nm.  The distance between layers,
  the rotation of the layers with respect to adjacent layers, and the number of
  monomers per layer do influence the equilibrium structure and require further
  justification for their choices. We rely on experimental data to inform our
  choices

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction.  A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. There will be no further discussion of this system but the interested
  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also unfeasible for the monomers to orient in the t-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  As outlined in \ref{point:monomernum} the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

  \subsection*{Equilibration}

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 1e6 KJ mol$^-1$ nm$^-2$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^-1$ nm$^-2$, the restraints are released linearly until there is no more
  restraining potential. The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run a long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time consant of 10 ps.
  In order to create a wet system, we solvated an initial configuration with
  water using gmx solvate. All water molecules placed outside the pore region are
  removed. Waters inside the pore region are randomly removed until the desired
  concentration of water in the pores is reached. The remainder of the
  equilibration follows the same procedure as the dry system. In all cases, the
  v-rescale thermostat was used with tau-t = 0.1. % Supplemental

  \subsection*{Crosslinking}

  Using an equilibrated structure, we performed a crosslinking procedure
  in order to match synthetic procedures. The purpose of crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
  Direction of attack was not considered because the resultant mixture is
  racemic.

  \subsection*{Equilibrium Calculations}

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is
  equilibrated based on when the distance between pores stops changing.

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. Pore centers are located by averaging the coordinates of sodium
  ions in their respective pores. Pore spacing statistics were generated 
  using the bootstrapping technique (See Supplemental Information).

  To quantify the degree of layering and the equilibrium distance between layers
  in our system, we calculate a spatial correlation function, $g(z)$, measured
  along the z-axis (perpendicular to the membrane plane). To calculate $g(z)$,
  we binned the z-component distances between the center of mass of each
  component and all others of the same pore over at least 50 ns of equilibrated
  trajectory and then normalized by the average number density. To extract the
  average distance between layers we applied a discrete fourier transform to
  $g(z)$ and extracted the highest intensity frequency.

  % BJC: To be replaced by Joe's description
  Simulated X-ray diffraction patterns are generated based on atomic
  coordinates in order to make a direct experimental comparison. All atomic
  coordinates were simulated as gaussian spheres of electron density
  corresponding to each atom's atomic number. A three dimensional fourier
  transform (FT) of the array of electron density results in a three dimensional
  structure factor which represents the unit cell in reciprocal space. We matched
  experimental 2D WAXS patterns by adjusting the initial spacing between layers
  and the orientation of the head groups with respect to adjacent layers.
